### 2021-10-26 papers/modeling-of-combined-lfr-csp-sco2-paper/content.tex
review edits on paper, word document in -> review response -> brian review is cover letter

Supercritical CO$_2$ Brayton cycles are promising cycle configurations offering higher efficiencies, compact design, and reduced turbomachinery cost while operating with non toxic working fluid. Various sCO$_2$ Brayton cycles have been modeled with the recompression cycle having efficiency advantages over other proposed cycle arrangements \cite{turchi_2013,ahn_2014,wang_2018}. The literature shows that the recompression cycle can reach efficiencies of 50\% in some scenarios (with turbine inlet temperatures in the 650-700 $^{\circ}$C range) \cite{turchi_2013,wright_2009}, allows these cycles to be a competitive alternative to steam Rankine and air Brayton cycles over a range of temperatures.   Due to the benefits of sCO$_2$ Brayton cycles, the United States Department of Energy is investigating these conversion cycles for use with heat sources including nuclear and solar \cite{doe_2012}. Multiple project funding opportunities are established with the National Energy Technology Laboratory offering 144 million dollar award for demonstration and performance verification of a sCO$_2$ Brayton cycle \cite{netl_2016} and the Office of Energy Efficiency \& Renewable Energy offering 2.6 million dollar reward in their Brayton Energy project for a integrated CSP receiver, TES, and power block in one sCO$_2$ Brayton system \cite{seto_2015}. Utilizing complementary technologies, specifically solar concentrating power and lead-cooled fast reactors, can offset the drawbacks of each. Coupling these technologies into an interconnected cycle allows for consistent generation, independent of weather or time of day, and thermal storage for high dispatchability during high grid demand periods.
A CSP has an array of mirrors concentrating solar rays towards a receiver to generate thermal energy therefore causing a dependency on weather conditions and time of day. Previous research has modeled CSP technology as the thermal source for Brayton sCO$_2$ cycles with promising results in efficiency gains and high temperature thermal energy storage \cite{turchi_2013, iverson_2013, ho_2015, wang_2018, neises_2020}. Thermal energy gathered from the CSP is stored in thermal energy storage which is dispatched into the sCO$_2$ Brayton cycle when grid demand increases. With a comparable temperature to the hot TES, 560 $^{\circ}$C, a LFR is capable of transferring heat to thermal storage for later dispatch. An LFR uses fission reactions to heat a heat transfer fluid, in this case liquid lead, to temperatures of 650 $^{\circ}$C. Nuclear sCO$_2$ Brayton cycles have been studied with similar gains in efficiency seen in CSP sCO$_2$ Brayton cycles \cite{ dostal_2004, luo_2020}. Physical testing of simple sCO$_2$ Brayton cycles has been performed at Sandia National Laboratories \cite{wright_2011} and in Korea \cite{cha_2016}, with at least two commercial companies (Echogen, Net Power) having progressed to building pilot cycles  \cite{held_2015,fetvedt_2016}.

@@

%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
%\begin{quote}
%This is an example of a quote.
%\end{quote}

@@

\subsection{Cycle Component Modeling}
%======================================================================================
Components present in the cycles are modeled using various techniques and are discussed in more detail below. Turbines and compressors are analyzed using isentropic efficiencies. Counter-flow heat exchangers are modeled using the effectiveness-NTU method while simplified ``black box'' heat exchangers that use a simple energy balance for state point calculations are used in lieu of more detailed component models where data is available. The lead-cooled fast reactor is assumed to be a black box heat exchanger because of the constant heat input and state points on the sCO$_2$ inlet and outlet are provided. The molten salt loop for the CSP is modeled with necessary components including hot and cold TES, receiver, pumps and counter-flow heat exchangers.

@@

\subsubsection{Black Box and Counter-Flow Heat Exchangers}
%--------------------------------------------------------------------------------------

Black box heat exchangers are simplified heat exchangers which have no approach temperature or pinch point and are modeled as a perfect heat transfer into or out of the cycle. These heat exchangers use an energy balance with mass flow inlet energy, heat input or output, and mass flow outlet energy. The energy balance equation used for all black box heat exchangers is Equation \ref{eq-black}.

@@

Counter-flow heat exchangers are modeled with two fluids flowing in opposite directions exchanging heat from the hot side to the cold side. The temperatures of the hot and cold flows on either side of the the heat exchanger have a temperature difference known as an approach temperature. A diagram showing a simplified counter-flow heat exchanger is illustrated in Figure \ref{counter-flow-hx}.

@@

Counter-flow heat exchanger calculations require two known state points, fluid properties, mass flow rate of hot and cold side, and a specified approach temperature. In the modeled cases, the approach temperature is set to value of $10^{\circ}$C, based off prior model development of sCO$_{2}$ Brayton cycle heat exchangers \cite{seidel_2010_model_development}. The fluid libraries referenced are built into EES for Carbon Dioxide and Salt (60\% NaNO$_3$ 40\% KNO$_3$) \cite{pacheco_1995_salt_properties,roland_1996_co2_properties}.

@@

Knowing the state points on all inlets and outlets of the counter-flow heat exchanger allows for the heat exchanger performance metrics to be calculated. Performance metrics include effectiveness, capacitance ratio, UA, and NTU for heat exchangers. Effectiveness is the ratio of the actual heat transfer rate to the maximum heat transfer rate, or a perfect heat exchanger with no approach temperature. Assuming the approach temperature is on the high side, the maximum heat transfer rate, $\dot{Q}_{max}$ is found with the maximum enthalpy. Maximum enthalpy of the cold stream is found with correlations by setting the temperature to T$_{in,h}$ with same pressure on the cold outlet. Using the maximum enthalpy, $h_{max}$, the maximum heat transfer rate is calculated using Equation \ref{heattrans_max}.

@@

Where $C_{avg}$ is the average capacitance rate across the hot or cold stream and $c_{in}$ and $c_{out}$ is the specific heat at the inlet and outlet respectively. Specific heat is found using library correlations. Once both average capacitances are calculated for the hot and cold streams, one has a larger value, $\dot{C}_{max}$, and one has a smaller value,  $\dot{C}_{min}$. These maximum and minimum values are used to find the capacitance ratio, $CR$, in Equation \ref{cap_ratio}.

@@

Where the left hand side, $\dot{m}$, $h_{inlet}$, and $\dot{Q}_{LFRHX}$, is the energy into the flow and the right hand side, $\dot{m}$ and $h_{outlet}$, is the energy brought out from the flow of sCO$_2$. The amount of energy transferred into the cycle, $\dot{Q}_{LFRHX}$, is set at $950$ MW, and outlet temperature of the sCO$_2$ from the LFR HX is set at a value of $595^{\circ}$C. The outlet temperature of the LFR is specified because of high temperature material limits on the LFR lead side. The low temperature side is allowed to vary over a range of values with some considerations. The lead flow velocity is limited by the erosion of the fuel, the slower the lead flow velocity reduces fuel erosion and therefore leads to a more desirable compact design. Restricting the lead flow velocity, and therefore lead mass flow rate, leads to a higher LFR power output when the inlet sCO$_2$ temperature is reduced. LFR sCO$_2$ inlet temperature has a lower bound of $340^{\circ}$C before the lead begins to freeze, which is operationally unacceptable. When the inlet temperature of sCO$_2$ is increased the temperature difference across the LFR is decreased leading to an increase in power conversion cycle thermodynamic efficiency but a reduction in LFR power below $950$ MW. There is a compromise between high LFR efficiency and LFR power, therefore a temperature of $400^{\circ}$C for the sCO$_2$ inlet temperature is the optimal value provided by Westinghouse.

@@

The CSP salt cycle modeled in this paper is composed of hot and cold thermal energy storage (TES), pumps, receiver, sCO$_2$-to-salt counter-flow heat exchanger (C2S), and CSP counter-flow heat exchanger (CSP HX). The diagram for this CSP salt loop is seen in Figure \ref{csp}.

@@

When grid demand for electrical power increases, a series of operating modes are activated. During the highest demand times, cycle operation focuses on maximum electrical generation. This is achieved through the C2S being turned off for direct electrical production from the LFR and the hot TES is discharging heat through the CSP HX for electrical production in the sCO$_2$ Brayton cycle. As grid demand diminishes, CSP HX ramps down heat extraction until no power is being dispatched through the salt and the hot TES begins charging. During this process, the LFR gradually adds a larger fraction of heat input to the TES through C2S. This process continues until no electrical production is occurring in the cycle and all heat is stored in TES for later use.

@@

The values displayed in Table \ref{tab-cycle-constants} are representative of CSP and the Westinghouse LFR.

In addition to standardized parameters, all cycles have identical recompression sides. The recompression side contains a precooler, low temperature recuperator, and two compressors; main compressor and recompressor. 
The modeled cycles are summarized in Table \ref{tab-cycle_sum}.

@@

In practice, sCO$_2$ cycles typically have 3 turbines, with 2 of these driving the compressor and recompressor. Therefore, this configuration will not in general require additional turbines compared to the C-1HTR1T-ON configuration. Furthermore, it is anticipated that the cost of the high temperature recuperator is likely most related to its volume, and hence having two smaller high temperature recuperators is unlikely to cost significantly more than one large one.

@@

For all cycle configurations presented, with constrained or unconstrained LFR low temperature inlet and varied cold TES temperature, cycle and heat storage efficiencies are maximized using parametric studies on flow splitter fractions. These results are obtained using standardized values found in Table \ref{tab-cycle_sum} for a more direct comparison between cycles. Presentation and discussion of the results is fulfilled in the following section.

@@

The calculated values reported in Table \ref{tab-noncharg} are component power, mass flow fractions, and counter-flow heat exchanger specifications. Columns labeled with '$C$' headers have the LFR sCO$_2$ low inlet temperature constrained to 400$^{\circ}$C while those labeled with '$U$' are unconstrained and calculated based off of a parametric study on the MC mass flow fraction. Values listed in the Cold TES Temperature row are set to values of 390$^{\circ}$C, 410$^{\circ}$C, or 440$^{\circ}$C to study the effects that cold TES temperature has on cycle efficiency and to accommodate certain characteristics of cycle configurations. Cycles that do not contain the listed component and omit the associated values. 
\clearpage 
% \mw{Check that the column breakdown and grouping is correct in what I've changed.}
\end{paracol}
\begin{specialtable}[H] 
    %[htbp]

>>>>

\replaced[comment={Abbreviation defined [4,1]}]{Supercritical carbon dioxide (sCO$_2$).}{Supercritical CO$_2$} Brayton cycles are promising cycle configurations offering higher efficiencies, compact design, and reduced turbomachinery cost while operating with non toxic working fluid. Various sCO$_2$ Brayton cycles have been modeled with the recompression cycle having efficiency advantages over other proposed cycle arrangements \cite{turchi_2013,ahn_2014,wang_2018}. The literature shows that the recompression cycle can reach efficiencies of 50\% in some scenarios (with turbine inlet temperatures in the 650-700 $^{\circ}$C range) \cite{turchi_2013,wright_2009}, allows these cycles to be a competitive alternative to steam Rankine and air Brayton cycles over a range of temperatures. Due to the benefits of sCO$_2$ Brayton cycles, the United States Department of Energy is investigating these conversion cycles for use with heat sources including nuclear and solar \cite{doe_2012}. Multiple project funding opportunities are established with the National Energy Technology Laboratory offering 144 million dollar award for demonstration and performance verification of a sCO$_2$ Brayton cycle \cite{netl_2016} and the Office of Energy Efficiency \& Renewable Energy offering 2.6 million dollar reward in their Brayton Energy project for a integrated \replaced[comment={Abbreviations defined [4,1]}]{concentrating solar power (CSP) receiver, thermal energy storage (TES),.}{CSP receiver, TES,} and power block in one sCO$_2$ Brayton system \cite{seto_2015}. Utilizing complementary technologies, specifically solar concentrating power and lead-cooled fast reactors, can offset the drawbacks of each. \replaced[comment={Drawbacks of CSP and LFR technology elaborated on [2,2]}]{These drawbacks include CSP dependency on weather conditions and time of day, while the LFR has a large system inertia on start-up and ramping while reducing strain on system components when consistently operating.}{} Coupling these technologies into an interconnected cycle allows for consistent generation, independent of weather or time of day, and thermal storage for high dispatchability during high grid demand periods. 
A CSP has an array of mirrors concentrating solar rays towards a \replaced[comment={Furthering explanation of the CSP [3,1]}]{tower receiver, transferring the energy into a solar salt, and storing the solar salt in tanks. CSP systems require direct sunlight to operate at full power therefore causing variability based on weather conditions and time of day.}{receiver to generate thermal energy therefore causing a dependency on weather conditions and time of day.} Previous research has modeled CSP technology as the thermal source for Brayton sCO$_2$ cycles with promising results in efficiency gains and high temperature thermal energy storage \cite{turchi_2013, iverson_2013, ho_2015, wang_2018, neises_2020}. Thermal energy gathered from the CSP is stored in thermal energy storage which is dispatched into the sCO$_2$ Brayton cycle when grid demand increases. With a comparable temperature to the hot TES, 560 $^{\circ}$C, a \replaced[comment={Abbreviation defined [4,1]}]{lead-cooled fast reactor (LFR).}{LFR} is capable of transferring heat to thermal storage for later dispatch. An LFR uses fission reactions to heat a heat transfer fluid, in this case liquid lead, to temperatures of 650 $^{\circ}$C. Nuclear sCO$_2$ Brayton cycles have been studied with similar gains in efficiency seen in CSP sCO$_2$ Brayton cycles \cite{ dostal_2004, luo_2020}. Physical testing of simple sCO$_2$ Brayton cycles has been performed at Sandia National Laboratories \cite{wright_2011} and in Korea \cite{cha_2016}, with at least two commercial companies (Echogen, Net Power) having progressed to building pilot cycles  \cite{held_2015,fetvedt_2016}.

@@

%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
%\begin{quote}
%This is an example of a quote.
%\end{quote}
%\replaced[comment={better}]{asdf}{test}

@@

\subsection{Cycle Component Modeling}
%======================================================================================
Components present in the cycles are modeled using various techniques and are discussed in more detail below. Turbines and compressors are analyzed using isentropic efficiencies. Counter-flow heat exchangers (HX) are modeled using the effectiveness-NTU method while simplified ``black box'' heat exchangers that use a simple energy balance for state point calculations are used in lieu of more detailed component models where data is available. The lead-cooled fast reactor is assumed to be a black box heat exchanger because of the constant heat input and state points on the sCO$_2$ inlet and outlet are provided. The molten salt loop for the CSP is modeled with necessary components including hot and cold TES, receiver, pumps and counter-flow heat exchangers.

@@

\subsubsection{Black Box and Counter-Flow Heat Exchangers}
%--------------------------------------------------------------------------------------
Black box heat exchangers are simplified heat exchangers which have no approach temperature or pinch point and are modeled as a perfect heat transfer into or out of the cycle. These heat exchangers use an energy balance with mass flow inlet energy, heat input or output, and mass flow outlet energy. The energy balance equation used for all black box heat exchangers is Equation \ref{eq-black}.

@@

Counter-flow heat exchangers are modeled with two fluids flowing in opposite directions exchanging heat from the hot side to the cold side. \replaced[comment={Improving counter-flow hx explanation [1,1]}]{The smallest temperature difference of the hot and cold flows on either the low or high end of the heat exchanger is defined as the approach temperature of the counter-flow heat exchanger.}{The temperatures of the hot and cold flows on either side of the heat exchanger have a temperature difference known as an approach temperature.} A diagram showing a simplified counter-flow heat exchanger is illustrated in Figure \ref{counter-flow-hx}.

@@

Counter-flow heat exchanger calculations require two known state points, fluid properties, mass flow rate of hot and cold side, and a specified approach temperature. In the modeled cases, the approach temperature is set to value of $10^{\circ}$C, based off prior model development of sCO$_{2}$ Brayton cycle heat exchangers \cite{seidel_2010_model_development}. The fluid libraries referenced are built into \replaced[comment={Abbreviation defined [4,1]}]{Engineering Equation Solver (EES)}{EES} for Carbon Dioxide and Salt (60\% NaNO$_3$ 40\% KNO$_3$) \cite{pacheco_1995_salt_properties,roland_1996_co2_properties}.

@@

Knowing the state points on all inlets and outlets of the counter-flow heat exchanger allows for the heat exchanger performance metrics to be calculated. Performance metrics include \replaced[comment={Abbreviations defined [4,1]}]{effectiveness ($\varepsilon$), capacitance (CR), conductivity (UA), and number of transfer units (NTU).}{effectiveness, capacitance, UA, and NTU} for heat exchangers. Effectiveness is the ratio of the actual heat transfer rate to the maximum heat transfer rate, \replaced[comment={sentence proofread}]{a perfect heat exchanger has an effectiveness of one}{or a perfect heat exchanger} with no approach temperature. Assuming the approach temperature is on the high side, the maximum heat transfer rate, $\dot{Q}_{max}$ is found with the maximum enthalpy. Maximum enthalpy of the cold stream is found with correlations by setting the temperature to T$_{in,h}$ with same pressure on the cold outlet. Using the maximum enthalpy, $h_{max}$, the maximum heat transfer rate is calculated using Equation \ref{heattrans_max}.

@@

Where $C_{avg}$ is the average capacitance rate across the hot or cold stream and $c_{in}$ and $c_{out}$ is the specific heat at the inlet and outlet respectively. \replaced[comment={Improving counter-flow hx eplanation [1,1]}]{Specific heat is found using library correlations, with the average capacitance rate assumed to be constant during the analysis. Assuming constant average capacitance rate is suitable for most engineering purposes, especially when there is uncertainty associated with other parameters \cite{nellis_klein_2008}.}{Specific heat is found using library correlations.} Once both average capacitances are calculated for the hot and cold streams, one has a larger value, $\dot{C}_{max}$, and one has a smaller value,  $\dot{C}_{min}$. These maximum and minimum values are used to find the capacitance ratio, $CR$, in Equation \ref{cap_ratio}.

@@

Where the left hand side, $\dot{m}$, $h_{inlet}$, and $\dot{Q}_{LFRHX}$, is the energy into the flow and the right hand side, $\dot{m}$ and $h_{outlet}$, is the energy brought out from the flow of sCO$_2$. The amount of energy transferred into the cycle, $\dot{Q}_{LFRHX}$, is set at $950$ MW, and outlet temperature of the sCO$_2$ from the LFR HX is set at a value of $595^{\circ}$C. The outlet temperature of the LFR is specified because of high temperature material limits on the LFR lead side. The low temperature side is allowed to vary over a range of values with some considerations. The lead flow velocity is limited by the erosion of the fuel, the slower the lead flow velocity reduces fuel erosion and therefore leads to a more desirable compact design. \replaced[comment={Edited sentence for clarity [2,3]}]{Restricting the lead flow velocity, and therefore lead mass flow rate, while lowering the inlet sCO$_2$ temperature increases the temperature difference across the LFR HX allowing for a higher, more desirable, LFR power output.}{Restricting the lead flow velocity, and therefore lead mass flow rate, leads to a higher LFR power output when the inlet sCO$_2$ temperature is reduced.} LFR sCO$_2$ inlet temperature has a lower bound of $340^{\circ}$C before the lead begins to freeze, which is operationally unacceptable. When the inlet temperature of sCO$_2$ is increased the temperature difference across the LFR is decreased leading to an increase in power conversion cycle thermodynamic efficiency but a reduction in LFR power below $950$ MW. There is a compromise between high LFR efficiency and LFR power, therefore a temperature of $400^{\circ}$C for the sCO$_2$ inlet temperature is the optimal value provided by Westinghouse.

@@

The CSP salt cycle modeled in this paper is composed of hot and cold thermal energy storage \replaced[comment={Removed Abbreviation [4,1]}]{}{(TES)}, pumps, receiver, sCO$_2$-to-salt counter-flow heat exchanger (C2S), and CSP counter-flow heat exchanger (CSP HX). The diagram for this CSP salt loop is seen in Figure \ref{csp}.

@@

When grid demand for electrical power increases, a series of operating modes are activated. During the highest demand times, cycle operation focuses on maximum electrical generation. This is achieved through the C2S being turned off for direct electrical production from the LFR and the hot TES is discharging heat through the CSP HX for electrical production in the sCO$_2$ Brayton cycle. \replaced[comment={Edited sentence, CSP still charges the hot TES during low grid demand periods [2,4]}]{As grid demand diminishes, electrical generation is ramped down by decreasing heat input through the CSP HX. This continues until no heat is being dispatched from the solar salt to the sCO$_2$, and all produced thermal energy from the CSP is put into charging the hot TES.}{As grid demand diminishes, CSP HX ramps down heat extraction until no power is being dispatched through the salt and the hot TES begins charging.} During this process, the LFR gradually adds a larger fraction of heat input to the TES through C2S. This process continues until no electrical production is occurring in the cycle and all heat is stored in TES for later use.

@@

The values displayed in Table \ref{tab-cycle-constants} are representative of CSP and the Westinghouse LFR. \replaced[comment={Elaborated on table variable names [2,5]}]{In Table \ref{tab-cycle-constants}, variable names require further explanation. The high side pressure with the variable label $P_{2A}$, is the constant pressure outlet on the compressors and inlet to the turbines. In all models the value is set by the outlet of the main compressor and held constant with the assumption that there is no pressure drop across heat exchangers. In addition to this pressure, temperatures are also set. The inlet temperature of the main compressor, $T_{1A}$, is set to a value of $40^{\circ}$C in all models. The temperature of the sCO$_2$ on the outlet of the LFR HX has different variable names, T$_5$, T$_{2C}$, T$_{6A}$, and T$_{5C}$, according to the associated cycle configuration diagram.}{}.

In addition to standardized parameters, all cycles have identical recompression sides. The recompression side contains a  \replaced[comment={Abbreviations defined [4,1]}]{precooler (PC), low temperature recuperator (LTR), and two compressors; main compressor (MC) and recompressor (RC)}{precooler, low temperature recuperator, and two compressors; main compressor and recompressor}. 
The modeled cycles are summarized in Table \ref{tab-cycle_sum}.

@@

In practice, sCO$_2$ cycles typically have 3 turbines, with 2 of these driving the compressor and recompressor. Therefore, this configuration will not in general require additional turbines compared to the C-1HTR1T-ON configuration. Furthermore, it is anticipated that the cost of the high temperature recuperator (HTR) is likely most related to its volume, and hence having two smaller high temperature recuperators is unlikely to cost significantly more than one large one.

@@

For all cycle configurations presented, with constrained or unconstrained LFR low temperature inlet and varied cold TES temperature, cycle and heat storage efficiencies are maximized using parametric studies on flow splitter fractions. \replaced[comment={Source of calculations defined [4,3]}]{All calculations were carried out using EES, with the results}{These results are} obtained using standardized values found in Table \ref{tab-cycle_sum} for a more direct comparison between cycles. Presentation and discussion of the results is fulfilled in the following section.

@@

The calculated values reported in Table \ref{tab-noncharg} are component power, mass flow fractions, and counter-flow heat exchanger specifications. Columns labeled with '$C$' headers have the LFR sCO$_2$ low inlet temperature constrained to 400$^{\circ}$C while those labeled with '$U$' are unconstrained and calculated based off of a parametric study on the MC mass flow fraction. Values listed in the Cold TES Temperature row are set to values of 390$^{\circ}$C, 410$^{\circ}$C, or 440$^{\circ}$C to study the effects that cold TES temperature has on cycle efficiency and to accommodate certain characteristics of cycle configurations. \replaced[comment={Removed the conjunction 'and' in the sentence for clarity. [1,1]}]{Cycles that do not contain the listed component omit the associated values.}{Cycles that do not contain the listed component and omit the associated values.} 
\clearpage 
% \mw{Check that the column breakdown and grouping is correct in what I've changed.}
\end{paracol}
\begin{specialtable}[H] 
    %[htbp]
